I don't know if projection is categorical. \\ Did \citealt{mandelkern-etal2020} discover that it is?

### 2023-11-17 writing/naturalness-paper/paper.tex
updated naturalness paper

\begin{table}[!h]
\addtolength{\tabcolsep}{-.19em}
\centering
\begin{tabular}{r | cccccccccccccccccccccccccc}
& \rots{continue} & \rots{too} & \rots{also} & \rots{again} & \rots{stop} & \rots{\color{orange}be annoyed\color{black}} & \rots{cleft} & \rots{\color{orange}know\color{black}} & \rots{\color{green}demonstrate\color{black}} & \rots{\color{green}pretend\color{black}} & \rots{\color{green}inform\color{black}} & \rots{\color{green}confess\color{black}} & \rots{\color{orange}see\color{black}} & \rots{\color{green}acknowledge\color{black}} & \rots{\color{orange}discover\color{black}} & \rots{\color{green}admit\color{black}} & \rots{\color{green}hear\color{black}} & \rots{\color{green}prove\color{black}} & \rots{\color{orange}reveal\color{black}} & \rots{\color{green}announce\color{black}} & \rots{\color{green}be right\color{black}} & \rots{\color{green}establish\color{black}} & \rots{\color{green}confirm\color{black}} & \rots{\color{green}suggest\color{black}} & \rots{\color{green}think\color{black}} & \rots{\color{green}say\color{black}} \\
\hline
\input{../../results/main/13explicitIgnorance/models/latex-tables/table1}
\hline
\end{tabular}
\caption{Posterior distributions of estimated marginal means of pairwise differences in the explicit ignorance context, with expressions ordered by mean naturalness rating in the same context. Color coding indicates whether the difference was positive or negative, and the size of the difference; a white cell means that there was no difference: \\
Positive difference: \colorbox{yellow1}{\makebox[4em][c]{$>=-1.5$}}, \colorbox{yellow2}{\makebox[4em][c]{($-1.5, -0.5]$}}, \colorbox{yellow3}{\makebox[4em][c]{($-0.5, 0]$}} \\
Negative difference: \colorbox{purple1}{\makebox[4em][c]{$>=1.5$}}, \colorbox{purple2}{\makebox[4em][c]{($1.5,0.5]$}}, 
\colorbox{purple3}{\makebox[4em][c]{($0.5, 0]$}} 
}\label{t:pairwise}
\end{table}

@@

\begin{table}[!h]
\addtolength{\tabcolsep}{-.19em}
\centering
\begin{tabular}{r  cccccccccccccccccccc}
& \rots{\color{orange}be annoyed\color{black}} & \rots{\color{orange}know\color{black}} & \rots{\color{green}demonstrate\color{black}} & \rots{\color{green}pretend\color{black}} & \rots{\color{green}inform\color{black}} & \rots{\color{green}confess\color{black}} & \rots{\color{orange}see\color{black}} & \rots{\color{green}acknowledge\color{black}} & \rots{\color{orange}discover\color{black}} & \rots{\color{green}admit\color{black}} & \rots{\color{green}hear\color{black}} & \rots{\color{green}prove\color{black}} & \rots{\color{orange}reveal\color{black}} & \rots{\color{green}announce\color{black}} & \rots{\color{green}be right\color{black}} & \rots{\color{green}establish\color{black}} & \rots{\color{green}confirm\color{black}} & \rots{\color{green}suggest\color{black}} & \rots{\color{green}think\color{black}} & \rots{\color{green}say\color{black}} \\
\hline
\input{../../results/main/13explicitIgnorance/models/latex-tables/table2}
\hline
\end{tabular}
\caption{Posterior distribution of estimated marginal means of pairwise differences between the three contexts for each expression, with expressions ordered by mean naturalness rating in explicit ignorance context. `EIC' stands for explicit ignorance context. Color coding indicates whether the difference was positive or negative, and the size of the difference; a white cell means that there was no difference: \\
Positive difference: \colorbox{yellow1}{\makebox[4em][c]{$>=-1.5$}}, \colorbox{yellow2}{\makebox[4em][c]{($-1.5, -0.5]$}}, \colorbox{yellow3}{\makebox[4em][c]{($-0.5, 0]$}} \\
Negative difference: \colorbox{purple1}{\makebox[4em][c]{$>=1.5$}}, \colorbox{purple2}{\makebox[4em][c]{($1.5,0.5]$}}, 
\colorbox{purple3}{\makebox[4em][c]{($0.5, 0]$}} 
}\label{t:pairwise2}
\end{table}

@@

This result has several methodological and theoretical implications. First, as the result aligns with the result of \citealt{degen-tonhauser-openmind} based on an inference rating measure, it suggests that inference rating measures and naturalness rating measures are sensitive to some of the same factors. Second, naturalness ratings in explicit ignorance and other contexts are sensitive not to whether content is presupposed or not (as claimed in \citealt{mandelkern-etal2020}) but to a whole slew of factors, including whether content is defeasible, at-issueness, prosody, and the prior probability of content. Finally, for analyses of presuppositions that assume that presuppositions can be locally accommodated, the result means that local accommodation must be sensitive to the prior probability of content, not just to whether the content is contradictory or uninformative in the context.

{\bf What do the results tell us about the lexical semantics of the different predicates?}

@@

What the question is not about is the projection of {\em utterance content}. What we mean by that is the strength of the projection inference that a particular interpreter draws from a particular utterance of a sentence in which a content is contributed in the scope of an entailment-canceling operator. The empirical evidence pointed to in \S\ref{s1} shows that different interpreters draw projection inferences of varying strength. For instance, as shown in Fig.~\ref{fig:dt1a-JULIAN}, the participants of \citepos{degen-tonhauser-language} Exp.~1a drew projection inferences of varying strength for the CC that Julian dances salsa from utterances of the interrogative {\em Did Scott discover that Julian dances salsa?} It is uncontroversial that this utterance content variability is gradient, not categorical. It is gradient because the projection of utterance content is sensitive to several factors on which interpreters differ, including their prior beliefs about the speaker or about the utterance content, their perception and interpretation of the prosodic realization of the utterances, and their assumptions about the question under discussion addressed by the utterance (e.g., \citealt{djaerv-bacovcin2020,degen-tonhauser-openmind,mahler2020,mahler-thesis,tonhauser-salt26,tbd-variability,tonhauser-etal-sub23}).\footnote{Variation can also be introduced by lso: noise, inattentiveness, but replication of variation exclude this. {\bf FIX THIS}}

@@

What the question about whether projection is gradient or categorical is about is {\em content}, where a content is, for instance, the CC of {\em discover}, the CC of {\em inform}, or the pre-state content of {\em stop}. The question, thus, is about whether projection is gradient or categorical when we abstract away from variation that is attributable to interpreter variability and consider the observed by-content variation (for instance that the CC of {\em discover} is less projective than the CC of {\em know}; see Fig.~\ref{fig:dt1a}): Is this by-content variation derived from underlying representations that either categorically trigger projection inferences or not, or from underlying representations that license projection inferences in combination with other factors in the uttered sentence?\footnote{Another potential source of projection variability is speaker belief. \citealt[498f.]{tbd-variability} pointed out that speaker belief could be categorical (that is, they either believe that is is true or they are uncertain) or gradient (in which case they believe that the content is true to some degree). On either view, interpreters might draw projection inferences of different strengths due to uncertainty about (the degree of) the speaker's belief.}

@@

So, where does this leave us with respect to the question of whether the projection of content is categorical or gradient? The challenge is clear: Proponents of either position must confront the observed by-content projection variability and develop analyses that capture the variability in a principled way. As none of the contemporary analyses of projection are able to do so, the verdict is still out.

@@

\section{Model output: Pairwise comparison of expressions}\label{a:analysis1}

{\bf JT PROPOSES TO LEAVE OUT THIS AND THE NEXT FIGURE}

Fig.~\ref{fig:comparisons1} provides a graphical representation of the results of the pairwise comparisons between the ratings for each expression in the explicit ignorance context. The full model output in table form is available here: \url{LINK TO TABLE IN REPO}.

\begin{sidewaysfigure}[h!]
\centering
\includegraphics[width=\textwidth]{../../results/main/13explicitIgnorance/graphs/comparisons-in-EIC}
\caption{.}\label{fig:comparisons1}
\end{sidewaysfigure}

\section{Model output: Pairwise comparison of contexts}\label{a:analysis2}

Fig.~\ref{fig:comparisons2} provides a graphical representation of the results of the pairwise comparisons of the ratings in the three contexts for each of the 20 predicates. The full model output in table form is available here: \url{LINK TO TABLE IN REPO}.

\begin{figure}[h!]
\centering
\includegraphics[width=.8\textwidth]{../../results/main/13explicitIgnorance/graphs/context-comparisons}
\caption{.}\label{fig:comparisons2}
\end{figure}

>>>>

\begin{table}[!h]
\addtolength{\tabcolsep}{-.19em}
\centering
\begin{tabular}{r | cccccccccccccccccccccccccc}
& \rots{continue} & \rots{too} & \rots{also} & \rots{again} & \rots{stop} & \rots{\color{orange}be annoyed\color{black}} & \rots{cleft} & \rots{\color{orange}know\color{black}} & \rots{\color{green}demonstrate\color{black}} & \rots{\color{green}pretend\color{black}} & \rots{\color{green}inform\color{black}} & \rots{\color{green}confess\color{black}} & \rots{\color{orange}see\color{black}} & \rots{\color{green}acknowledge\color{black}} & \rots{\color{orange}discover\color{black}} & \rots{\color{green}admit\color{black}} & \rots{\color{green}hear\color{black}} & \rots{\color{green}prove\color{black}} & \rots{\color{orange}reveal\color{black}} & \rots{\color{green}announce\color{black}} & \rots{\color{green}be right\color{black}} & \rots{\color{green}establish\color{black}} & \rots{\color{green}confirm\color{black}} & \rots{\color{green}suggest\color{black}} & \rots{\color{green}think\color{black}} & \rots{\color{green}say\color{black}} \\
\hline
\input{../../results/main/13explicitIgnorance/models/analysis1/table1}
\hline
\end{tabular}
\caption{Posterior distributions of estimated marginal means of pairwise differences in the explicit ignorance context, with expressions ordered by mean naturalness rating in the same context. Color coding indicates whether the difference was positive or negative, and the size of the difference; a white cell means that there was no difference: \\
Positive difference: \colorbox{yellow1}{\makebox[4em][c]{$>=-1.5$}}, \colorbox{yellow2}{\makebox[4em][c]{($-1.5, -0.5]$}}, \colorbox{yellow3}{\makebox[4em][c]{($-0.5, 0]$}} \\
Negative difference: \colorbox{purple1}{\makebox[4em][c]{$>=1.5$}}, \colorbox{purple2}{\makebox[4em][c]{($1.5,0.5]$}}, 
\colorbox{purple3}{\makebox[4em][c]{($0.5, 0]$}} 
}\label{t:pairwise}
\end{table}

@@

\begin{table}[!h]
\addtolength{\tabcolsep}{-.19em}
\centering
\begin{tabular}{r  cccccccccccccccccccc}
& \rots{\color{orange}be annoyed\color{black}} & \rots{\color{orange}know\color{black}} & \rots{\color{green}demonstrate\color{black}} & \rots{\color{green}pretend\color{black}} & \rots{\color{green}inform\color{black}} & \rots{\color{green}confess\color{black}} & \rots{\color{orange}see\color{black}} & \rots{\color{green}acknowledge\color{black}} & \rots{\color{orange}discover\color{black}} & \rots{\color{green}admit\color{black}} & \rots{\color{green}hear\color{black}} & \rots{\color{green}prove\color{black}} & \rots{\color{orange}reveal\color{black}} & \rots{\color{green}announce\color{black}} & \rots{\color{green}be right\color{black}} & \rots{\color{green}establish\color{black}} & \rots{\color{green}confirm\color{black}} & \rots{\color{green}suggest\color{black}} & \rots{\color{green}think\color{black}} & \rots{\color{green}say\color{black}} \\
\hline
\input{../../results/main/13explicitIgnorance/models/analysis2/table2}
\hline
\end{tabular}
\caption{Posterior distribution of estimated marginal means of pairwise differences between the three contexts for each expression, with expressions ordered by mean naturalness rating in explicit ignorance context. `EIC' stands for explicit ignorance context. Color coding indicates whether the difference was positive or negative, and the size of the difference; a white cell means that there was no difference: \\
Positive difference: \colorbox{yellow1}{\makebox[4em][c]{$>=-1.5$}}, \colorbox{yellow2}{\makebox[4em][c]{($-1.5, -0.5]$}}, \colorbox{yellow3}{\makebox[4em][c]{($-0.5, 0]$}} \\
Negative difference: \colorbox{purple1}{\makebox[4em][c]{$>=1.5$}}, \colorbox{purple2}{\makebox[4em][c]{($1.5,0.5]$}}, 
\colorbox{purple3}{\makebox[4em][c]{($0.5, 0]$}} 
}\label{t:pairwise2}
\end{table}

@@

This result has several methodological and theoretical implications. First, as the result aligns with the result of \citealt{degen-tonhauser-openmind} based on an inference rating measure, it suggests that inference rating and naturalness rating measures are sensitive to some of the same factors. Second, naturalness ratings in explicit ignorance and other contexts are sensitive to a whole slew of factors, including the (implicit) prosody of the utterance and the defeasibility, at-issueness, and the prior probability of content. Finally, for analyses of presuppositions that local accommodation, the result means that local accommodation is sensitive to the prior probability of content, not just to whether the content is contradictory or uninformative in the context.

%One thing that is striking about the naturalness ratings is that for many  predicates the ratings are quite high in both the explicit ignorance and the higher prior probability context (e.g., {\em inform, see, discover, hear, reveal, be right, think, say}). This suggests that interrogatives with these predicates are natural both in contexts in which the speaker is uncertain about the truth of the CC and in contexts in which the speaker believes the CC to have a high probability. This result, which crosscuts the factive/nonfactive distinction, highlights the flexibility with which these predicates are used in discourse,

@@

What the question is not about is the projection of {\em utterance content}. What we mean by that is the strength of the projection inference that a particular interpreter draws from a particular utterance of a sentence in which a content is contributed in the scope of an entailment-canceling operator. The empirical evidence pointed to in \S\ref{s1} shows that different interpreters draw projection inferences of varying strength. For instance, as shown in Fig.~\ref{fig:dt1a-JULIAN}, the participants of \citepos{degen-tonhauser-language} Exp.~1a drew projection inferences of varying strength for the CC that Julian dances salsa from utterances of the interrogative {\em Did Scott discover that Julian dances salsa?} It is uncontroversial that this utterance content variability is gradient, not categorical. It is gradient because the projection of utterance content is sensitive to several factors on which interpreters differ, including their prior beliefs about the speaker or about the utterance content, their perception and interpretation of the prosodic realization of the utterances, and their assumptions about the question under discussion addressed by the utterance (e.g., \citealt{djaerv-bacovcin2020,degen-tonhauser-openmind,mahler2020,mahler-thesis,tonhauser-salt26,tbd-variability,tonhauser-etal-sub23}).

@@

What the question about whether projection is gradient or categorical is about is {\em content}, where a content is, for instance, the CC of {\em discover}, the CC of {\em inform}, or the pre-state content of {\em stop}. The question, thus, is about whether projection is gradient or categorical when we abstract away from variation that is attributable to interpreter variability and consider the observed by-content variation (for instance that the CC of {\em discover} is less projective than the CC of {\em know}; see Fig.~\ref{fig:dt1a}): Is this by-content variation derived from underlying representations that either categorically trigger projection inferences or not, or from underlying representations that license projection inferences in combination with other factors in the uttered sentence?\footnote{Another potential source of projection variability is speaker belief. \citealt[498f.]{tbd-variability} pointed out that speaker belief could be categorical (the speaker either believe that the content is true or they are uncertain) or gradient (the speaker believes that the content is true to some degree). On either view, interpreters might draw projection inferences of different strengths due to uncertainty about (the degree of) the speaker's belief.}

@@

So, where does this leave us with respect to the question of whether the projection of content is categorical or gradient? The challenge is clear: Proponents of either position must confront the observed by-content projection variability and develop analyses that capture the variability in a principled way. As none of the contemporary projection analyses are able to do so, one might say that the verdict is still out. The empirical evidence, however, seems to favor the hypothesis that projection is gradient, as there is, to date, no empirical evidence that projection is categorical, and several pieces of evidence that it is gradient.

@@

%\section{Model output: Pairwise comparison of expressions}\label{a:analysis1}
%
%{\bf JT PROPOSES TO LEAVE OUT THIS AND THE NEXT FIGURE}
%
%Fig.~\ref{fig:comparisons1} provides a graphical representation of the results of the pairwise comparisons between the ratings for each expression in the explicit ignorance context. The full model output in table form is available here: \url{LINK TO TABLE IN REPO}.
%
%\begin{sidewaysfigure}[h!]
%\centering
%\includegraphics[width=\textwidth]{../../results/main/13explicitIgnorance/graphs/comparisons-in-EIC}
%\caption{.}\label{fig:comparisons1}
%\end{sidewaysfigure}
%
%\section{Model output: Pairwise comparison of contexts}\label{a:analysis2}
%
%Fig.~\ref{fig:comparisons2} provides a graphical representation of the results of the pairwise comparisons of the ratings in the three contexts for each of the 20 predicates. The full model output in table form is available here: \url{LINK TO TABLE IN REPO}.
%
%\begin{figure}[h!]
%\centering
%\includegraphics[width=.8\textwidth]{../../results/main/13explicitIgnorance/graphs/context-comparisons}
%\caption{.}\label{fig:comparisons2}
%\end{figure}
